# The Basics for R

Before you can ask your computer to save some numbers, you’ll need to know how to talk to it. That’s where R and RStudio come in. RStudio gives you a way to talk to your computer. R gives you a language to speak in. 

*R* is a cross-platform, zero-cost statistical programming environment that offers a powerful and easy-to-learn set of statistical/machine learning tools to assist you with finding the insights in your own data. Here we cover the fundamental skills for the statistical programming language *R*.

## R

To download R, go to CRAN, the comprehensive R archive network. CRAN is composed of a set of mirror servers distributed around the world and is used to distribute R and R packages. Don’t try and pick a mirror that’s close to you: instead use the cloud mirror, <https://cloud.r-project.org>, which automatically figures it out for you.

A new major version of R comes out once a year, and there are 2-3 minor releases each year. It’s a good idea to update regularly. Upgrading can be a bit of a hassle, especially for major versions, which require you to re-install all your packages, but putting it off only makes it worse.

## RStudio

RStudio is an integrated development environment, or IDE, for R programming. Download and install it from <http://www.rstudio.com/download>. RStudio is updated a couple of times a year. When a new version is available, RStudio will let you know. It’s a good idea to upgrade regularly so you can take advantage of the latest and greatest features. For this book, make sure you have at least RStudio 1.0.0.

When you start RStudio, you’ll see two key regions in the interface:


For now, all you need to know is that you type R code in the console pane, and press enter to run it. You’ll learn more as we go along!

## The liver package

You'll also need to install some R packages.
An R **package** is a collection of functions, data, and documentation that extends the capabilities of base R.
Using packages is key to the successful use of R.
The majority of the packages that you will learn in this book are part of the so-called tidyverse.
All packages in the tidyverse share a common philosophy of data and R programming, and are designed to work together naturally.

You can install the complete tidyverse with a single line of code:

```{r eval = FALSE}
install.packages( "liver" )
```

On your own computer, type that line of code in the console, and then press enter to run it.
R will download the packages from CRAN and install them on to your computer.
If you have problems installing, make sure that you are connected to the internet, and that <https://cloud.r-project.org/> isn't blocked by your firewall or proxy.

You will not be able to use the functions, objects, or help files in a package until you load it with `library()`.
Once you have installed a package, you can load it using the `library()` function:

```{r message=FALSE}
library( liver )
```

This tells you that tidyverse is loading eight packages: ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr, and forcats packages.
These are considered to be the **core** of the tidyverse because you'll use them in almost every analysis.






